Comparing Hydrogen and Jet-A for an Ultra High Bypass Ratio Turbofan with Water Recirculation

### 2023-04-27 conference-paper/paper_main.tex

% Update with table from Andrew
% TODO: AL - PA, What is the citation key "Kalnay1996"?  It's not in the MDOLab Bib or references.bib
\begin{table}[hbt!]
    \centering
    \caption{
        N3 engine model parameters.
        These parameters specify the engine design used for analysis in this report.
        Atmospheric parameters are presented that specify the flight conditions and humidity of the atmosphere.}
    \begin{tabular}{c c c c}
        \hline
        Parameter             & Value    & Units                 & Comments                                   \\
        \hline
        $W_{TOC}$             & 820.441  & $lbm/s$               & TOC total mass flow rate of air            \\
        $BPR_{TOC}$           & 23.945   & $-$                   & TOC bypass Ratio                           \\
        $PR_{HPC,TOC}$        & 53.633   & $-$                   & TOC overall pressure ratio                 \\
        $PR_{LPC,TOC}$        & 3.00     & $-$                   & TOC low-pressure compressor pressure ratio \\
        $PR_{fan,TOC}$        & 1.30     & $-$                   & TOC fan pressure ratio                     \\
        $F_{net,SLS}$         & 28620.84 & $lbf$                 & SLS net thrust                             \\
        $F_{net,CRZ}$         & 5510.72  & $lbf$                 & CRZ net thrust                             \\
        $T_{4,TOC}/T_{4,RTO}$ & 0.926    & $-$                   & TOC-to-RTO temperature ratio               \\
        $T_{4,RTO}$           & 3400.0   & $^{\circ}R$           & RTO combustor temperature                  \\
        $h_{TOC}$             & 0.00017  & $kg_{water}/kg_{air}$ & humidity ratio at TOC \cite{Kalnay1996}    \\
        $h_{CRZ}$             & 0.00017  & $kg_{water}/kg_{air}$ & humidity ratio at CRZ \cite{Kalnay1996}    \\
        $h_{RTO}$             & 0.009    & $kg_{water}/kg_{air}$ & humidity ratio at RTO \cite{Kalnay1996}    \\
        $h_{SLS}$             & 0.009    & $kg_{water}/kg_{air}$ & humidity ratio at SLS \cite{Kalnay1996}    \\
        $LHV_{JetA}$          & 18564.0  & $BTU/lbm$             & Lower heating value of Jet-A               \\
        $LHV_{H2}$            & 51591.0  & $BTU/lbm$             & Lower heating value of H2                  \\
        \hline
    \end{tabular}
    \label{engine_params}
\end{table}

@@

This optimization problem was run for both Jet-A and H2 fuels using the parameters given in Table \ref{engine_params}.
The optimization problem for each fuel was specified as:

\begin{table}[h]
    \centering
    \caption{
        Multipoint optimization problem definition.
        The 4 design variables are the water recovery fractions of the core exhaust stream at each operation condition.
        The objective function is the thrust specific energy consumption at the cruise condition with a net thrust constraint at the design point, TOC.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l}
        \toprule
                        & Variable/Function           & Description                                 & Units           & Quantity \\
        \hline
        minimize        & $ \dot{m}_{fuel,CRZ} $      & Fuel flow rate at CRZ                       & $\frac{lbm}{s}$ & 1        \\
                        &                             &                                             &                 &          \\
        with respect to & $x_{H2O,TOC}$               & Water recovery fraction at TOC              & -               & 1        \\
                        & $x_{H2O,RTO}$               & Water recovery fraction at RTO              & -               & 1        \\
                        & $x_{H2O,SLS}$               & Water recovery fraction at SLS              & -               & 1        \\
                        & $x_{H2O,CRZ}$               & Water recovery fraction at CRZ              & -               & 1        \\
                        & $PR_{HPC,TOC}$              & TOC high-pressure compressor pressure ratio & -               & 1        \\
                        & $PR_{LPC,TOC}$              & TOC low-pressure compressor pressure ratio  & -               & 1        \\
                        & $PR_{fan,TOC}$              & TOC fan pressure ratio                      & -               & 1        \\
                        & $T_{4,TOC}/T_{4,RTO}$       & TOC-to-RTO temperature ratio                & -               & 1        \\
                        & $T_{4,RTO}$                 & RTO combustor temperature                   & $^\circ$R       & 1        \\
                        & $V_{ratio,CRZ}$             & RTO combustor temperature                   & $^\circ$R       & 1        \\
        \cline{3-5}
                        &                             & Total                                       &                 & 4        \\
                        &                             &                                             &                 &          \\
        subject to      & $F_{net,TOC} \geq 5800$ lbf & Target net thrust at TOC                    & lbf             & 1        \\
                        & $D_{Fan} \leq 100$ $in^2$   & Maximum Fan Diameter                        & $in^2$          & 1        \\
        \cline{3-5}
                        &                             & Total                                       &                 & 1        \\
        \bottomrule
    \end{tabular}
    \label{tab:opt_problem}
\end{table}

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{JetA-H2_bar_chart_diff.pdf}
    \caption{Percent relative improvement in thrust specific energy consumption (TSEC) of an ultra-high bypass turbofan engine from water injection.
        Data was collected with and without water injected into the core stream of the engine for Jet-A and hydrogen fuels.}
    \label{fig:results}
\end{figure}

>>>>

% Update with table from Andrew
% TODO: AL - PA, What is the citation key "Kalnay1996"?  It's not in the MDOLab Bib or references.bib
\begin{table}[hbt!]
    \centering
    \caption{
        N+3 engine model parameters.
        These parameters specify the engine design used for analysis in this report.
        Atmospheric parameters are presented that specify the flight conditions and humidity of the atmosphere.}
    \begin{tabular}{c c c c}
        \hline
        Parameter             & Value    & Units                 & Comments                                   \\
        \hline
        $W_{TOC}$             & 820.441  & $lbm/s$               & TOC total mass flow rate of air            \\
        $BPR_{TOC}$           & 23.945   & $-$                   & TOC bypass Ratio                           \\
        $PR_{HPC,TOC}$        & 53.633   & $-$                   & TOC overall pressure ratio                 \\
        $PR_{LPC,TOC}$        & 3.00     & $-$                   & TOC low-pressure compressor pressure ratio \\
        $PR_{fan,TOC}$        & 1.30     & $-$                   & TOC fan pressure ratio                     \\
        $F_{net,SLS}$         & 28620.84 & $lbf$                 & SLS net thrust                             \\
        $F_{net,CRZ}$         & 5510.72  & $lbf$                 & CRZ net thrust                             \\
        $T_{4,TOC}/T_{4,RTO}$ & 0.926    & $-$                   & TOC-to-RTO temperature ratio               \\
        $T_{4,RTO}$           & 3400.0   & $^{\circ}R$           & RTO combustor temperature                  \\
        $h_{TOC}$             & 0.00017  & $kg_{water}/kg_{air}$ & humidity ratio at TOC \cite{Kalnay1996}    \\
        $h_{CRZ}$             & 0.00017  & $kg_{water}/kg_{air}$ & humidity ratio at CRZ \cite{Kalnay1996}    \\
        $h_{RTO}$             & 0.009    & $kg_{water}/kg_{air}$ & humidity ratio at RTO \cite{Kalnay1996}    \\
        $h_{SLS}$             & 0.009    & $kg_{water}/kg_{air}$ & humidity ratio at SLS \cite{Kalnay1996}    \\
        $LHV_{JetA}$          & 18564.0  & $BTU/lbm$             & Lower heating value of Jet-A               \\
        $LHV_{H2}$            & 51591.0  & $BTU/lbm$             & Lower heating value of H2                  \\
        \hline
    \end{tabular}
    \label{engine_params}
\end{table}

@@

This optimization problem was run for both Jet-A and H2 fuels using the parameters given in Table \ref{engine_params}.
% The optimization problem for each fuel was specified as:

\begin{table}[h]
    \centering
    \caption{
        Multipoint optimization problem definition.
        The 4 design variables are the water recovery fractions of the core exhaust stream at each operation condition.
        The objective function is the fuel flow rate at the CRZ condition with a net thrust and fan diameter constraint at the design point, TOC.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l}
        \toprule
                        & Variable/Function           & Description                                & Units           & Quantity \\
        \hline
        minimize        & $ \dot{m}_{fuel,CRZ} $      & Fuel flow rate at CRZ                      & $\frac{lbm}{s}$ & 1        \\
                        &                             &                                            &                 &          \\
        with respect to & $x_{H2O,TOC}$               & Water recovery fraction at TOC             & -               & 1        \\
                        & $x_{H2O,RTO}$               & Water recovery fraction at RTO             & -               & 1        \\
                        & $x_{H2O,SLS}$               & Water recovery fraction at SLS             & -               & 1        \\
                        & $x_{H2O,CRZ}$               & Water recovery fraction at CRZ             & -               & 1        \\
                        & $PR_{fan,TOC}$              & TOC fan pressure ratio                     & -               & 1        \\
                        & $PR_{LPC,TOC}$              & TOC low-pressure compressor pressure ratio & -               & 1        \\
                        & $PR_{OPR,TOC}$              & TOC overall pressure ratio                 & -               & 1        \\
                        & $T_{4,TOC}/T_{4,RTO}$       & TOC-to-RTO temperature ratio               & -               & 1        \\
                        & $T_{4,RTO}$                 & RTO combustor temperature                  & $^\circ$R       & 1        \\
                        & $V_{ratio,CRZ}$             & Velocity exit ratio at CRZ                 &                 & 1        \\
        \cline{3-5}
                        &                             & Total                                      &                 & 4        \\
                        &                             &                                            &                 &          \\
        subject to      & $F_{net,TOC} \geq 5800$ lbf & Target net thrust at TOC                   & lbf             & 1        \\
                        & $D_{Fan} \leq 100$ $in^2$   & Maximum Fan Diameter                       & $in^2$          & 1        \\
        \cline{3-5}
                        &                             & Total                                      &                 & 1        \\
        \bottomrule
    \end{tabular}
    \label{tab:opt_problem}
\end{table}

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{JetA-H2_TSEC_diff.pdf}
    \caption{Percent relative improvement in thrust specific energy consumption (TSEC) of an ultra-high bypass turbofan engine from water injection.
        Data was collected with and without water injected into the core stream of the engine for Jet-A and hydrogen fuels.}
    \label{fig:barchart}
\end{figure}
